---
output:
  github_document:
    html_preview: false
---

```{r readme-01, echo = FALSE}
knitr::opts_chunk$set(
  fig.asp = 2/3,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# photobiologyLEDs

[![CRAN version](https://www.r-pkg.org/badges/version-last-release/photobiologyLEDs)](https://cran.r-project.org/package=photobiologyLEDs) 
[![cran checks](https://cranchecks.info/badges/worst/photobiologyLEDs)](https://cran.r-project.org/web/checks/check_results_photobiologyLEDs.html)

Package '**photobiologyLEDs**' contains spectral emission data for diverse types of light emitting diodes (LEDs) and LED arrays. It complements other packages in the suite of R packages for photobiology 'r4photobiology'. This package contains only data.
 
## Examples

```{r example-02, message=FALSE}
library(photobiologyLEDs)
```

How many spectra are included in the current version of 'photobiologyFilters'?

```{r example-03}
length(leds.mspct)
```

What are the names of available spectra. We use `head()` to limit the output.

```{r example-04}
# list names of the first 10 LEDs
head(names(leds.mspct), 10)
```

To subset based on different criteria we can use predefined character vectors of LED names. For example, vector `nichia` lists the names of the spectra for filters made by NIchia. 

```{r example-05}
nichia
```

We can use the vector to extract all these spectra as a collection.

```{r example-05a}
leds.mspct[nichia]
```

The package includes a character vector with the names of these vectors of names that are available.

```{r example-05b}
all_led_accessors
```

Please, see the _User Guide_ or documentation
for the names of other vectors of names for materials, suppliers, and regions of
the spectrum.

Summary calculations can be easily done with methods from package 'photobiology'. Here we calculate mean transmittance for two regions of the spectrum given by wavelengths in nanometres.

```{r example-06}
q_irrad(leds.mspct[["NS6L183AT_H1"]]) * 1e6
```

The `autoplot()` methods from package 'ggspectra' can be used for plotting one or more spectra at a time. The classes of the objects used to store the spectral data are derived from `"data.frame"` making direct use of the data easy with functions and methods 
from base R and various packages.

## Installation

Installation of the most recent stable version from CRAN:

```{r, eval=FALSE}
install.packages("photobiologyLEDs")
```

Installation of the current unstable version from Bitbucket:

```{r, eval=FALSE}
# install.packages("devtools")
devtools::install_bitbucket("aphalo/photobiologyleds")
```

## Documentation 

HTML documentation is available at (https://docs.r4photobiology.info/photobiologyLEDs/), including a _User Guide_.

News on updates to the different packages of the 'r4photobiology' suite are regularly posted at (https://www.r4photobiology.info/).

Two articles introduce the basic ideas behind the design of the suite and its use:
Aphalo P. J. (2015) (https://doi.org/10.19232/uv4pb.2015.1.14) and Aphalo P. J. (2016)  (https://doi.org/10.19232/uv4pb.2016.1.15).

A book is under preparation, and the draft is currently available at (https://leanpub.com/r4photobiology/).

A handbook written before the suite was developed contains useful information on the quantification and manipulation of ultraviolet and visible radiation:
Aphalo, P. J., Albert, A., Björn, L. O., McLeod, A. R., Robson, T. M., & Rosenqvist, E. (Eds.) (2012) Beyond the Visible: A handbook of best practice in plant UV photobiology (1st ed., p. xxx + 174). Helsinki: University of Helsinki, Department of Biosciences, Division of Plant Biology. ISBN 978-952-10-8363-1 (PDF), 978-952-10-8362-4 (paperback). PDF file available from (https://hdl.handle.net/10138/37558).

## Contributing

Pull requests, bug reports, and feature requests are welcome at (https://bitbucket.org/aphalo/photobiologyfilters).

## Citation

If you use this package to produce scientific or commercial publications, please cite according to:

```{r}
citation("photobiologyLEDs")
```

## License

© 2012-2020 Pedro J. Aphalo (pedro.aphalo@helsinki.fi). Released under the GPL, version 2 or greater. This software carries no warranty of any kind.
