---
title: "`photobiologyLEDs` `r packageVersion('photobiologyLEDs')`<br>Data Catalogue"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data Catalogue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
knitr::opts_chunk$set(fig.width=8, fig.height=4)
```


## Introduction

This is a catalogue of all data sets included in the package. You will find below plots of the emission spectra of different lamps. All spectra are normalized to an area of one at the peak of the curve.

The spectral data have been acquired mostly with one instrument. However, some
spectra haver been measured with a spectrometer that has lower wavelength 
resolution giving for the same leds measured peaks of slightly different width. 
This is an inevitable artefact of spectral measurements, but as LEDs have 
relatively wide peaks the distortion is small.

```{r}
library(photobiology)
library(photobiologyWavebands)
library(photobiologyLEDs)
library(ggplot2)
library(ggspectra)
```


```{r}
options(photobiology.plot.annotations =
          c("boxes", "labels", "colour.guide", "peaks", "title"))
```

## All LEDs in data set

```{r}
names(leds.mspct)
```

## UVB LEDs

```{r}
plot(leds.mspct$TY_UV310nm)
plot(leds.mspct$UVMAX305)
```

## UVA LEDs

```{r}
plot(leds.mspct$UVMAX340)
plot(leds.mspct$G_P30R140A1_XT)
plot(leds.mspct$XSL365)
plot(leds.mspct$XSL370)
plot(leds.mspct$UV395)
```


## VIS LEDs

### Violet and blue LEDs

```{r}
plot(leds.mspct$LED405)
plot(leds.mspct$LED435_66_60)
plot(leds.mspct$HPR40E_48K30BG)
plot(leds.mspct$CB30)
plot(leds.mspct$HLMB_CB30)
plot(leds.mspct$HLMP_CB31)
```

### Green and amber LEDs

```{r}
plot(leds.mspct$HLMP_CM30)
plot(leds.mspct$LY5436)
plot(leds.mspct$HLMP_DL32)
```

### Red and far-red LEDs

```{r}
plot(leds.mspct$HLMP_DJ32)
plot(leds.mspct$BS436)
plot(leds.mspct$B5_436_30D)
plot(leds.mspct$QDDH66002)
plot(leds.mspct$QDDH68002)
plot(leds.mspct$QDDH70002)
plot(leds.mspct$LED740_01AV)
plot(leds.mspct$QDDH73502)
plot(leds.mspct$FR_OLD)
```

## Broad band LEDs

```{r}
plot(leds.mspct$white)
plot(leds.mspct$Q36_4000K)
```

## Multichannel LEDs

```{r}
plot(leds.mspct$NHXRGB090_R)
plot(leds.mspct$NHXRGB090_G)
plot(leds.mspct$NHXRGB090_B)
```

```{r}
ggplot(leds.mspct$NHXRGB090) + 
  aes(color = channel) +
  geom_line() +
  labs(x = "Wavelength (nm)", y = "Normalized spectral irradiance",
       title = "leds.mspct$NHXRGB090")
```

