---
title: "`photobiologyLEDs` `r packageVersion('photobiologyLEDs')`<br>User Guide"
author: "Pedro J. Aphalo"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r}
knitr::opts_chunk$set(fig.width=8, fig.height=4)
```

```{r}
library(photobiology)
library(photobiologyWavebands)
library(photobiologyLEDs)
```

## Introduction

In this very brief User Guide we describe how to re-scale the normalized
spectra, and how to access individual spectra or subsets of spectra.

## Accessing individual spectra

The `source_spct` member objects in `leds.mspct` can be accessed through their
names or through a numeric index. As the numeric indexes are likely to change 
with updates to the package, their use is discouraged. Names as character
strings should be used instead. The names are listed in the documentation
and also available through the "Data Catalogue" vignette. They can also be
listed with method `names()`.

```{r}
names(leds.mspct)
```

We can use a character string as index to extract an individual `source_spct`
object.

```{r}
leds.mspct$UV395
```

```{r}
leds.mspct[["UV395"]]
```

Be aware that according to R's rules, using single square brackets will return
a `source_mspct` object possibly of length one. This statement is not equivalent
to the one in the chunk immediately above.

```{r}
leds.mspct["UV395"]
```

## Accessing subsets of spectra

We can subset the `source_mspct` object by indexing with vectors of character
strings. The package provides some predefined ones, and users can easily
define their own, either as constants or through computation. Here we use
a vector defined by the package.

```{r}
leds.mspct[norlux]
```

And below we use a computed one. In this case we extract the member spectra
with names containing the string "QDDH".

```{r}
leds.mspct[grep("QDDH", names(leds.mspct))]
```

## Calculating summaries from the normalized data

The spectra are normalized, and consequently, several summaries expressed in
absolute units are undefined, and trigger errors. Summaries like ratios which
are not affected by normalization are allowed and valid. The data have been
normalized as the measuring conditions used are not all the same, and in many
cases not well characterized (e.g. distance to nearby reflecting walls, or exact 
alignment of the spectrometer input optics with respect to light sources).

What we will do in this section is to rescale the spectral data so that after 
conversion a given target value for a summary quantity will be true. As an 
example, we will rescale one spectrum so that it yields an energy irradiance of
100 W m-2 for the range 315 to 400 nm.

```{r}
my.spct <- fscale(leds.mspct$UV395,
                  range = c(315, 400),
                  e_irrad,
                  target = 10
                  )
e_irrad(my.spct, waveband(c(315,400)))
```

If we want to treat the rescaled spectral data, as if they were true readings
with no scaling we can reset the attribute with method `setScaled()`. With 
method `getScaled()` we can test if a spectrum has been scaled.

```{r}
getScaled(my.spct)
setScaled(my.spct)
getScaled(my.spct)
```

If for some obscure reason we want to simply "pretend" that the spectral data
have not been normalized, we can permanently override the attribute on a copy 
of the data. Most of the time this is a very bad idea!

```{r}
my.UV395 <- leds.mspct$UV395
setNormalized(my.UV395)
e_irrad(my.UV395)
```

As mentioned above, ratios can be calculated directly as they are not affected
by normalization.

```{r}
q_ratio(leds.mspct$UV395, UVB(), UVA())
```
## Using the data in other contexts

As `source_spct` is a class derived from `list`, and `source_spct` is derived from
`tibble::tible` which is a compatible reimplementation of `data.frame` the
data can be used very easily with any R function.

```{r}
head(as.data.frame(leds.mspct$UV395))
```

Of course `attach` and `with` also work as expected.
```{r}
attach(leds.mspct)
q_ratio(UV395, UVB(), UVA())
detach(leds.mspct)
```

```{r}
attach(leds.mspct)
with(UV395, max(w.length))
detach(leds.mspct)
```

```{r}
with(leds.mspct, q_ratio(UV395, UVB(), UVA()))
```




